A Linear Solution of the Linearity Challenge

### 2025-12-23 paper/calculus.tex
cleanup
revised intro

\section{\Calculus}
\label{sec:calculus}

@@

\begin{table}[t]
    \caption{Syntax of \Calculus.}\label{tab:syntax}
    \begin{tabular}{c}
        \begin{math}
            \displaystyle
            \begin{array}[t]{rcll}
                P, Q & ::= & \Link\x\y            & \text{link} \\
                    & | & \Fail\x                 & \text{fail} \\
                    & | & \Wait\x.P               & \text{wait} \\
                    & | & \Close\x                & \text{close} \\
                    & | & \Case\x\z{P}{Q}         & \text{case} \\
                    & | & \Select\x{\InTag_i}\z.P & \text{left/right selection} \\
                    & | & \Join\x\y\z.P           & \text{join} \\
                    & | & \Fork\x\y\z{P}{Q}       & \text{fork} \\
                    & | & \Forall\x\X\y.P         & \text{for all} \\
                    & | & \Exists\x{A}\y.P        & \text{exists} \\
                    & | & \Server\x\y.P           & \text{server} \\
                    & | & \Client\x\y.P           & \text{client} \\
                    & | & \Weaken\x.P             & \text{weakening} \\
                    & | & \Contract\x\y\z.P       & \text{contraction} \\
                    & | & \Cut[A]\x{P}{Q}         & \text{cut} \\
            \end{array}
        \end{math}
    \end{tabular}
\end{table}

The syntax of \Calculus makes use of an infinite set of \emph{channel names},
ranged over by $x$, $y$ and $z$, and is shown in~\Cref{tab:syntax}.
%
A link $\Link\x\y$ denotes the merging of the channels $x$ and $y$, so that
every message sent on one of the channels is forwarded to the other. As
discussed in the literature~\cite{LindleyMorris15}, this form is also useful for
modeling \emph{delegation}, that is the exchange of an existing session endpoint
in another session.
%
The $\Wait\x.P$ and $\Close\x$ processes respectively model the input and output
of a termination signal on the channel $x$. The latter process has no
continuation, while the former process continues as $P$ once the signal has been
received.
%
The process $\Fail\x$ can be used to denote a failing protocol concerning the
channel $x$.
%
The processes $\Case\x\z{Q_1}{Q_2}$ and $\Select\x{\InTag_i}\z.P$ for $i=1,2$
describe the input/output of a label $\InTag_i$ along with a fresh continuation
$z$ on the channel $x$. The receiver continues as $Q_i$ depending on the
received label $\InTag_i$. In practice it is useful to consider arbitrary
(domain-specific) sets of tags instead of the canonical $\InTag_1$ and
$\InTag_2$. This generalization has no substantial impact on the formalization,
so we stick to the classical case of ``binary choices'' in line with the usual
presentations of linear logic.
%
The processes $\Join\x\y\z.R$ and $\Fork\x\y\z{P}{Q}$ describe the input/output
of a pair of fresh channels $y$ and $z$ on the channel $x$. The former process
forks into $P$ and $Q$, each using $y$ and $z$ respectively. The latter process
continues as $R$. These forms are often used in conjunction with links to model
\emph{delegation}.
%
The processes $\Exists\x{A}\y.P$ and $\Forall\x\X\y.P$ describe the input/output
of a type on the channel $x$ along with a fresh continuation $y$. These forms
are useful to describe polymorphic processes.

@@

\begin{table}[t]
    \caption{Operational semantics of \Calculus.}
    \label{tab:semantics}
    \begin{tabular}{@{}c@{}}
        \begin{math}
            \displaystyle
            \begin{array}{rrcll}
            \SLink &
            \Link\x\y & \pcong & \Link\y\x
            \\
            \SComm &
            \Cut\x{P}{Q} & \pcong & \Cut\x{Q}{P}
            \\
            \SFail &
            \Cut\x{\Fail\y}{P} & \pcong & \Fail\y % & x \ne y
            \\
            \SWait &
            \Cut\x{\Wait\y.P}{Q} & \pcong & \Wait\y.\Cut\x{P}{Q} % & x \ne y
            \\
            \SCase &
            \Cut\x{\Case\y\z{P}{Q}}{R} & \pcong & \Case\y\z{\Cut\x{P}{R}}{\Cut\x{Q}{R}} % & x \ne y,z
            \\
            \SSelect &
            \Cut\x{\Select\y{\InTag_i}\z.P}{Q} & \pcong & \Select\y{\InTag_i}\z.\Cut\x{P}{Q} % & x \ne y,z
            \\
            \SJoin &
            \Cut\x{\Join\y\u\v.P}{Q} & \pcong & \Join\y\u\v.\Cut\x{P}{Q} % & x \ne y,u,v
            \\
            \SForkL &
            \Cut\x{\Fork\y\u\v{P}{Q}}{R} & \pcong & \Fork\y\u\v{\Cut\x{P}{R}}{Q} & x \in \fn{P}
            \\
            \SForkR &
            \Cut\x{\Fork\y\u\v{P}{Q}}{R} & \pcong & \Fork\y\u\v{P}{\Cut\x{Q}{R}} & x \in \fn{Q} %\setminus\set{y,u,v}
            \\
            \SForall &
            \Cut\x{\Forall\y\X\z.P}{Q} & \pcong & \Forall\y\X\z.\Cut\x{P}{Q} % & x \ne y,z
            \\
            \SExists &
            \Cut\x{\Exists\y{A}\z.P}{Q} & \pcong & \Exists\y{A}\z.\Cut\x{P}{Q} % & x \ne y,z
            \\
            \SServer &
            \Cut\x{\Server\y\u.P}{\Server\x\v.Q} & \pcong & \Server\y\u.\Cut\x{P}{\Server\x\v.Q} % & x \ne y,u
            \\
            \SClient &
            \Cut\x{\Client\y\z.P}{Q} & \pcong & \Client\y\z.\Cut\x{P}{Q} % & x \ne y,z
            \\
            \SWeaken &
            \Cut\x{\Weaken\y.P}{Q} & \pcong & \Weaken\y.\Cut\x{P}{Q} % & x \ne y
            \\
            \SContract &
            \Cut\x{\Contract\y\u\v.P}{Q} & \pcong & \Contract\y\u\v.\Cut\x{P}{Q} % & x \ne y,u,v
            \\
            \\
            \RLink &
            \Cut\x{\Link\x\y}{P} & \red & P\subst\y\x \\
            \RClose &
            \Cut\x{\Close\x}{\Wait\x.P} & \red & P \\
            \RSelect &
            \Cut\x{\Select\x{\InTag_i}\z.P}{\Case\x\z{Q_1}{Q_2}} & \red & \Cut\z{P}{Q_i} & i\in\set{1,2} \\
            \RFork &
            \Cut\x{\Fork\x\y\z{P}{Q}}{\Join\x\y\z.R} & \red & \Cut\y{P}{\Cut\z{Q}{R}} \\
            \RExists &
            \Cut\x{\Exists\x{A}\z.P}{\Forall\x\X\z.Q} & \red &
            \Cut\z{P}{Q\subst{A}\X} \\
            \RConnect &
            \Cut\x{\Server\x\y.P}{\Client\x\y.Q} & \red & \Cut\y{P}{Q} \\
            \RWeaken &
            \Cut\x{\Server\x\y.P}{\Weaken\x.Q} & \red & \Weaken{\seqof\z}.Q
            & \fn{P} = \set{y,\seqof\z} \\
            \RContract &
            \Cut\x{\Server\x\y.P}{\Contract\x{x'}{x''}.Q} & \red &
            \multicolumn{2}{l}{
                \Contract{\seqof\z}{\seqof\z'}{\seqof\z''}.\Cut{x'}{
                    \Server{x'}\y.P'
                }{
                    \Cut{x''}{
                        \Server{x''}\y.P''
                    }{
                        Q
                    }
                }
                ~^*
            }
            \\
            \RCut &
            \Cut\x{P}{R} & \red & \Cut\x{Q}{R} & P \red Q \\
            \RCong &
            P & \red & Q & P \pcong R \red Q \\
            \\
            \multicolumn{5}{c}{
                ~^*
                \fn{P} = \set{y,\seqof\z},
                P' = P\subst{\seqof\z'}{\seqof\z},
                P'' = P\subst{\seqof\z''}{\seqof\z}
            }
            \end{array}
        \end{math}
    \end{tabular}
\end{table}

>>>>

\section{Linear \CP}
\label{sec:calculus}

@@

\begin{table}[t]
    \caption{Syntax of \Calculus.}\label{tab:syntax}
    \begin{tabular}{c}
        \begin{math}
            \displaystyle
            \begin{array}[t]{rcll}
                P, Q & ::= & \Link\x\y            & \text{link} \\
                    & | & \Fail\x                 & \text{fail} \\
                    & | & \Wait\x.P               & \text{wait} \\
                    & | & \Close\x                & \text{close} \\
                    & | & \Case\x\z{P}{Q}         & \text{case} \\
                    & | & \Select\x{\InTag_i}\z.P & \text{left/right selection} \\
                    & | & \Join\x\y\z.P           & \text{join} \\
                    & | & \Fork\x\y\z{P}{Q}       & \text{fork} \\
                    & | & \Forall\x\X\z.P         & \text{for all} \\
                    & | & \Exists\x{A}\z.P        & \text{exists} \\
                    & | & \Server\x\y.P           & \text{server} \\
                    & | & \Client\x\y.P           & \text{client} \\
                    & | & \Weaken\x.P             & \text{weakening} \\
                    & | & \Contract\x\y\z.P       & \text{contraction} \\
                    & | & \Cut[A]\x{P}{Q}         & \text{cut} \\
            \end{array}
        \end{math}
    \end{tabular}
\end{table}

The syntax of \Calculus makes use of an infinite set of \emph{channel names},
ranged over by $x$, $y$ and $z$, and is shown in~\Cref{tab:syntax}.
%
A link $\Link\x\y$ denotes the merging of the channels $x$ and $y$, so that
every message sent on one of the channels is forwarded to the other. As
discussed in the literature~\cite{LindleyMorris15}, this form is also useful for
modeling \emph{delegation}, that is the exchange of an existing session endpoint
in another session.
%
The $\Wait\x.P$ and $\Close\x$ processes respectively model the input and output
of a termination signal on the channel $x$. The latter process has no
continuation, while the former process continues as $P$ once the signal has been
received.
%
The process $\Fail\x$ can be used to denote a failing protocol concerning the
channel $x$.
%
The processes $\Case\x\z{Q_1}{Q_2}$ and $\Select\x{\InTag_i}\z.P$ for $i=1,2$
describe the input/output of a label $\InTag_i$ along with a fresh continuation
$z$ on the channel $x$. The receiver continues as $Q_i$ depending on the
received label $\InTag_i$. In practice it is useful to consider arbitrary
(domain-specific) sets of tags instead of the canonical $\InTag_1$ and
$\InTag_2$. This generalization has no substantial impact on the formalization,
so we stick to the classical case of ``binary choices'' in line with the usual
presentations of linear logic.
%
The processes $\Join\x\y\z.R$ and $\Fork\x\y\z{P}{Q}$ describe the input/output
of a pair of fresh channels $y$ and $z$ on the channel $x$. The former process
forks into $P$ and $Q$, each using $y$ and $z$ respectively. The latter process
continues as $R$. These forms are often used in conjunction with links to model
\emph{delegation}.
%
The processes $\Exists\x{A}\z.P$ and $\Forall\x\X\z.P$ describe the input/output
of a type on the channel $x$ along with a fresh continuation $z$. These forms
are useful to describe polymorphic processes.

@@

\begin{table}[t]
    \caption{Operational semantics of \Calculus.}
    \label{tab:semantics}
    \begin{tabular}{@{}c@{}}
        \begin{math}
            \displaystyle
            \begin{array}{rrcll}
            \SLink &
            \Link\x\y & \pcong & \Link\y\x
            \\
            \SComm &
            \Cut\x{P}{Q} & \pcong & \Cut\x{Q}{P}
            \\
            \SFail &
            \Cut\x{\Fail\y}{P} & \pcong & \Fail\y % & x \ne y
            \\
            \SWait &
            \Cut\x{\Wait\y.P}{Q} & \pcong & \Wait\y.\Cut\x{P}{Q} % & x \ne y
            \\
            \SCase &
            \Cut\x{\Case\y\z{P}{Q}}{R} & \pcong & \Case\y\z{\Cut\x{P}{R}}{\Cut\x{Q}{R}} % & x \ne y,z
            \\
            \SSelect &
            \Cut\x{\Select\y{\InTag_i}\z.P}{Q} & \pcong & \Select\y{\InTag_i}\z.\Cut\x{P}{Q} % & x \ne y,z
            \\
            \SJoin &
            \Cut\x{\Join\y\u\z.P}{Q} & \pcong & \Join\y\u\z.\Cut\x{P}{Q} % & x \ne y,u,v
            \\
            \SForkL &
            \Cut\x{\Fork\y\u\z{P}{Q}}{R} & \pcong & \Fork\y\u\z{\Cut\x{P}{R}}{Q} & x \in \fn{P}
            \\
            \SForkR &
            \Cut\x{\Fork\y\u\z{P}{Q}}{R} & \pcong & \Fork\y\u\z{P}{\Cut\x{Q}{R}} & x \in \fn{Q} %\setminus\set{y,u,v}
            \\
            \SForall &
            \Cut\x{\Forall\y\X\z.P}{Q} & \pcong & \Forall\y\X\z.\Cut\x{P}{Q} % & x \ne y,z
            \\
            \SExists &
            \Cut\x{\Exists\y{A}\z.P}{Q} & \pcong & \Exists\y{A}\z.\Cut\x{P}{Q} % & x \ne y,z
            \\
            \SServer &
            \Cut\x{\Server\y\u.P}{\Server\x\v.Q} & \pcong & \Server\y\u.\Cut\x{P}{\Server\x\v.Q} % & x \ne y,u
            \\
            \SClient &
            \Cut\x{\Client\y\z.P}{Q} & \pcong & \Client\y\z.\Cut\x{P}{Q} % & x \ne y,z
            \\
            \SWeaken &
            \Cut\x{\Weaken\y.P}{Q} & \pcong & \Weaken\y.\Cut\x{P}{Q} % & x \ne y
            \\
            \SContract &
            \Cut\x{\Contract\y\u\v.P}{Q} & \pcong & \Contract\y\u\v.\Cut\x{P}{Q} % & x \ne y,u,v
            \\
            \\
            \RLink &
            \Cut\x{\Link\x\y}{P} & \red & P\subst\y\x \\
            \RClose &
            \Cut\x{\Close\x}{\Wait\x.P} & \red & P \\
            \RSelect &
            \Cut\x{\Select\x{\InTag_i}\z.P}{\Case\x\z{Q_1}{Q_2}} & \red & \Cut\z{P}{Q_i} & i\in\set{1,2} \\
            \RFork &
            \Cut\x{\Fork\x\y\z{P}{Q}}{\Join\x\y\z.R} & \red & \Cut\y{P}{\Cut\z{Q}{R}} \\
            \RExists &
            \Cut\x{\Exists\x{A}\z.P}{\Forall\x\X\z.Q} & \red &
            \Cut\z{P}{Q\subst{A}\X} \\
            \RConnect &
            \Cut\x{\Server\x\y.P}{\Client\x\y.Q} & \red & \Cut\y{P}{Q} \\
            \RWeaken &
            \Cut\x{\Server\x\y.P}{\Weaken\x.Q} & \red & \Weaken{\seqof\z}.Q
            & \fn{P} = \set{y,\seqof\z} \\
            \RContract &
            \Cut\x{\Server\x\y.P}{\Contract\x{x'}{x''}.Q} & \red &
            \multicolumn{2}{l}{
                \Contract{\seqof\z}{\seqof\z'}{\seqof\z''}.\Cut{x'}{
                    \Server{x'}\y.P'
                }{
                    \Cut{x''}{
                        \Server{x''}\y.P''
                    }{
                        Q
                    }
                }
                ~^*
            }
            \\
            \RCut &
            \Cut\x{P}{R} & \red & \Cut\x{Q}{R} & P \red Q \\
            \RCong &
            P & \red & Q & P \pcong R \red Q \\
            \\
            \multicolumn{5}{c}{
                ~^*
                \fn{P} = \set{y,\seqof\z},
                P' = P\subst{\seqof\z'}{\seqof\z},
                P'' = P\subst{\seqof\z''}{\seqof\z}
            }
            \end{array}
        \end{math}
    \end{tabular}
\end{table}

### 2025-12-23 paper/introduction.tex
revised intro

One of the challenges in the benchmark, henceforth called \emph{linearity
challenge}, concerns the formalization of a \emph{session calculus}.
%
Sessions and session types~\citep{Honda93,HondaVasconcelosKubo98,HuttelEtAl16}
are established formalisms for the static analysis of distributed programs based
on peer-to-peer communication channels. Every session type system revolves
around three key ideas: (1) session endpoints are \emph{linear resources} that
cannot be discarded or duplicated without compromising some safety and liveness
properties of a program; (2) the type of a session endpoint is \emph{updated}
after each use to reflect the state of the interaction it describes; (3) peer
session endpoints are meant to be used in complementary ways so as to guarantee
the absence of communication errors and, to some extent, progress of the
interaction.
%
The linearity challenge is motivated by the observation that the proper handling
of linear resources in a formalization often requires a large number of complex
auxiliary definitions and technical results that divert from the main problem
under consideration. One of the alleged culprits of such unwanted complexity is
\emph{context splitting}, namely the operation that partitions a typing context
in such a way that the linear resources described therein end up in only one of
the partitions. This observation has led to the development of various
alternatives to context splitting including leftover
typing~\cite{ZalakainDardha21}, the use of linearity
predicates~\cite{SanoKavanaghPientka23} and tagged contexts~\cite{ZackonEtAl25}.

This paper proposes a logical and linear solution of the linearity challenge. By
``logical and linear solution'' we mean that the calculus that we formalize,
dubbed \Calculus, and the one in the challenge description differ for two key
aspects.
%
First of all, the challenge considers an \emph{ad hoc}, minimal calculus of
first-order sessions, whereas \LCP is a fully-featured calculus that supports
higher-order, polymorphic, linear and shared channels. Also, \LCP is closely
related to \CP~\citep{Wadler14} and its type system coincides with the proof
system of classical linear logic.
%
The second difference is that the calculus in the challenge is equipped with
\emph{native} sessions, whereas \LCP features linear channels that are meant to
be used for a \emph{single} communication. Structured sessions involving the
exchange of several messages can be modeled using \emph{explicit continuations}
according to the well-known encoding of (binary) sessions into linear
channels~\citep{KobayashiPierceTurner99,Kobayashi02b,DardhaGiachinoSangiorgi17}.
%
Overall, \LCP is a lower-level calculus in which sessions are encodable instead
of being featured natively, but it has the same expressive power of \CP.

We value these differences between \LCP and the calculus in the challenge
description because we think that they contribute to the simplicity of the
formalization more than the adopted formalization techniques, which end up being
fairly standard.\Luca{Our work fits the spirit of the challenge identifying not
only the most appropriate techniques but also the most appropriate models that
aid the formalization of concurrent/distributed models of computation.}
%
The fact of working with a calculus based on linear logic prevents \emph{by
construction} the same (sequential) process to own both endpoints of a session.
This situation is clearly undesirable since it does not correspond to a useful
pattern of interaction (every meaningful session requires its endpoints to be
used by parallel processes) and is a potential source of deadlocks.
%
From the standpoint of the formalization, where the representation of channel
names is a primary design choice, it spares us the need to \emph{distinguish}
the two endpoints of a session, \eg by means of polarities~\citep{GayHole05} or
by using different names connected by the same
binder~\citep{Vasconcelos12,GayVasconcelos25}.
%
Also, the fact of working with a calculus with \emph{linear} channels and
explicit continuations spares us the need to \emph{update} the type of channels
in typing contexts. When a channel is used it is effectively consumed, therefore
its type can be \emph{removed} from the typing context and the type of the
continuation channel is \emph{added} back to the typing context. As is turns
out, removing and adding types is easier than updating them. Even more so
considering that the type of continuation channels can be conveniently added
\emph{at the beginning} of a typing context, since such channels are fresh by
definition.

% Compared to the other analogous formalizations of session/linear
% calculi~\cite{GotoEtAl16,Thiemann19,CicconePadovani20,CastroFerreiraYoshida20,RouvoetPoulsenKrebbersVisser20,ZalakainDardha21,JacobsBalzerKrebbers22,SanoKavanaghPientka23,ZackonEtAl25},
% the one we present here is both the most complete (in terms of features
% supported by the calculus) and the most streamlined (in terms of code size). It
% is also one of only two formalizations that prove the deadlock freedom property
% for a session calculus, and it does so with substantially less code than the
% existing formalization~\cite{JacobsBalzerKrebbers22}.

\subparagraph*{Structure of the paper.}
%
\Cref{sec:calculus} describes the syntax and the operational semantics of \LCP
and states the properties of well-typed processes that we formalize and prove,
namely typing preservation, deadlock freedom and runtime safety.
%
\Cref{sec:agda} presents a roadmap of the Agda formalization and discusses its
key aspects with particular emphasis on the representation of channel names and
of typing contexts.
%
\Cref{sec:examples} illustrates a few examples of \LCP processes as they are
represented in Agda and shows how to use Agda as an interpreter for \LCP.
%
\Cref{sec:related} discusses related work more in detail by providing a
qualitative and quantitative comparison between the known formalizations of
session/linear calculi, including the one for \LCP.
%
\Cref{sec:conclusion} summarizes our contributions and discusses some ongoing
and future work.

>>>>

One of the challenges in the benchmark, henceforth called \emph{linearity
challenge}, concerns the formalization of a \emph{minimal calculus of sessions}.
%
Sessions and session types~\citep{Honda93,HondaVasconcelosKubo98,HuttelEtAl16}
are established formalisms for the static analysis of distributed programs based
on peer-to-peer communication channels. Every session type system revolves
around three key ideas: (1) session endpoints are \emph{linear resources} that
cannot be discarded or duplicated without compromising some safety and liveness
properties of a program; (2) the type of a session endpoint is \emph{updated}
after each use to reflect the state of the interaction it describes; (3) peer
session endpoints are meant to be used in complementary ways so as to guarantee
the absence of communication errors and, to some extent, progress of the
interaction.
%
The linearity challenge is motivated by the observation that the proper handling
of linear resources in a formalization often requires a large number of
auxiliary definitions and technical results that divert from the main problem
under study~\cite{CarboneEtAl24}. One of the alleged culprits of such complexity
is \emph{context splitting}, namely the operation that partitions a typing
context in such a way that the linear resources described therein end up in only
one of the partitions. This observation has led to the exploration of various
alternatives techniques including leftover typing~\cite{ZalakainDardha21}, the
use of linearity predicates~\cite{SanoKavanaghPientka23} and tagged
contexts~\cite{ZackonEtAl25}.

In this paper we attack the linearity challenge from a different perspective:
\emph{instead of describing new techniques that are effective for the
formalization of the calculus in the challenge, we describe a new calculus that
is easy to formalize with known techniques}.\Luca{dobbiamo dire perché lo
spirito della challenge è rispettato, e dunque giustificare il titolo
dell'articolo. Dire che il nostro non è un calcolo ad hoc, ma ha fondamenti
logici importanti, è completo, ed è strettamente in relazione con altri modelli
analoghi (CP e linear pi)}
%
More specifically, we present a calculus called \LCP whose type system coincides
with the proof system of \emph{classical linear logic} and that features
\emph{linear channels} instead of sessions. Linear channels are meant to be used
for a \emph{single} communication. Structured sessions involving the exchange of
several messages can be modeled in \LCP using \emph{explicit continuations}
according to the well-known encoding of (binary) sessions into the linear
$\pi$-calculus~\citep{KobayashiPierceTurner99,Kobayashi02b,DardhaGiachinoSangiorgi17}.
%
Overall, \LCP is a low-level calculus with a strong connection with linear logic
and in which sessions are encodable instead of being featured natively.
%
These features play an important role in taming the complexity of its
formalization.
%
The fact of working with a calculus based on linear logic prevents \emph{by
construction} the same (sequential) process to own both endpoints of a session.
This situation is clearly undesirable since it does not correspond to a useful
pattern of interaction (every meaningful session requires its endpoints to be
used by parallel processes) and is a potential source of deadlocks.
%
From the standpoint of the formalization, where the representation of channel
names is a primary design choice, it spares us the need to \emph{distinguish}
the two endpoints of a session, \eg by means of polarities~\citep{GayHole05} or
by using different names connected by the same
binder~\citep{Vasconcelos12,GayVasconcelos25}.
%
Also, the fact of working with a calculus with \emph{linear} channels and
explicit continuations spares us the need to \emph{update} the type of channels
in typing contexts. When a channel is used it is effectively consumed, therefore
its type can be \emph{removed} from the typing context and the type of the
continuation channel is \emph{added} back to the typing context. As is turns
out, removing and adding types is easier than updating them. Even more so
considering that the type of continuation channels can be conveniently added
\emph{at the beginning} of a typing context, since such channels are fresh by
definition.

In the end, the formalization of \LCP is both the most complete (in terms of
features supported by the calculus) and the most streamlined (in terms of code
size) among the known formalizations of session/linear
calculi~\cite{GotoEtAl16,Thiemann19,CastroFerreiraYoshida20,CicconePadovani20,RouvoetPoulsenKrebbersVisser20,SanoKavanaghPientka23,JacobsBalzerKrebbers22,ZackonEtAl25,ZalakainDardha21}.
It is also one of only two formalizations that prove the deadlock freedom
property for a session calculus, and it does so with substantially less code
than the existing formalization~\cite{JacobsBalzerKrebbers22}.

The rest of the paper is organized as follows.
%
\Cref{sec:calculus} describes the syntax and the operational semantics of \LCP
and states the properties of well-typed processes that we formalize and prove,
namely typing preservation, deadlock freedom and runtime safety.
%
\Cref{sec:agda} presents a roadmap of the Agda formalization and discusses its
key aspects with particular emphasis on the representation of channel names and
of typing contexts.
%
\Cref{sec:examples} illustrates a few examples of \LCP processes as they are
represented in Agda and shows how to use Agda as an interpreter for \LCP.
%
\Cref{sec:related} discusses related work more in detail by providing a
qualitative and quantitative comparison between the known formalizations of
session/linear calculi, including the one for \LCP.
%
\Cref{sec:conclusion} summarizes our contributions and discusses some ongoing
and future work.
